---
title: "ML_TESTING"
output: html_document
date: "2025-04-23"
---


```{r}
library(terra) 
library(raster)
library(tidyverse)
library(dplyr)
library(caret)
library(sp)
library(randomForest)
library(ggplot2)
library(sf)
library(rnaturalearth)
library(mgcv)  
library(gamlss)       
library(gamlss.add)   
library(maxnet)     
library(PresenceAbsence)  
library(dlnm)
library(pROC)
```

Need to layer tiff files to and extract trial and test data 
```{r}
getwd()

#Load TIFF files
tiff_file_2b_noirr_pic_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/Ethiopia_Wheat_noirrigation_5km_2020 copy.tif")  

tiff_file_2b_irr_pic_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/Ethiopia_Wheat_Irrigation_5km_2020 copy.tif")

tiff_file_2b_noirr_rcp26_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/Ethiopia_wheat_RCP26_noirrigation_5km_2020 copy.tif")
  
tiff_file_2b_irr_rcp26_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/Ethiopia_Wheat_RCP26_Irrigation_5km_2020 copy.tif")

tiff_file_2b_noirr_rcp60_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/Ethiopia_wheat_RCP60_noirrigation_5km_2020 copy.tif")

tiff_file_2b_irr_rcp60_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/Ethiopia_Wheat_RCP60_Irrigation_5km_2020 copy.tif")


#Check the number of layers and metadata
print(tiff_file_2b_noirr_pic_20)
plot(tiff_file_2b_noirr_pic_20)  

print(tiff_file_2b_irr_pic_20)
plot(tiff_file_2b_irr_pic_20)  

print(tiff_file_2b_noirr_rcp26_20)
plot(tiff_file_2b_noirr_rcp26_20)  

print(tiff_file_2b_irr_rcp26_20)
plot(tiff_file_2b_irr_rcp26_20)  

print(tiff_file_2b_noirr_rcp60_20)
plot(tiff_file_2b_noirr_rcp60_20) 

print(tiff_file_2b_irr_rcp60_20)
plot(tiff_file_2b_irr_rcp60_20) 
```

Load soil moisture files:
```{r}

#Load TIFF files
tiff_file_2b_soilmoist_pic_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/CLM45_PIC_Weighted_Average_Soil_Moisture_5km_2020 copy.tif")  

tiff_file_2b_soilmoist_rcp26_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/CLM45_RPC_26_Weighted_Average_Soil_Moisture_5km_2020 copy.tif")
  
tiff_file_2b_soilmoist_rcp60_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/CLM45_RPC_60_Weighted_Average_Soil_Moisture_5km_2020 copy.tif")

tiff_file_2b_soilmoist_rcp85_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/CLM45_RPC_85_Weighted_Average_Soil_Moisture_5km_2020 copy.tif")


#Check the number of layers and metadata
print(tiff_file_2b_soilmoist_pic_20)
plot(tiff_file_2b_soilmoist_pic_20)  
 
print(tiff_file_2b_soilmoist_rcp26_20)
plot(tiff_file_2b_soilmoist_rcp26_20)  

print(tiff_file_2b_soilmoist_rcp60_20)
plot(tiff_file_2b_soilmoist_rcp60_20) 

print(tiff_file_2b_soilmoist_rcp85_20)
plot(tiff_file_2b_soilmoist_rcp85_20) 
```


Load Food security value file:
```{r}
Food_security_values = read.csv("Socio-economic-survey-ETH_2020_FSI.csv")

View(Food_security_values)
```

```{r}

#Load TIFF files

#Yeild
tiff_file_2b_noirr_rcp26_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/Ethiopia_wheat_RCP26_noirrigation_5km_2020 copy.tif") 

print(tiff_file_2b_noirr_rcp26_20)
plot(tiff_file_2b_noirr_rcp26_20)  

#soil moisture
tiff_file_2b_soilmoist_rcp26_20 <- rast("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/CLM45_RPC_26_Weighted_Average_Soil_Moisture_5km_2020 copy.tif")

print(tiff_file_2b_soilmoist_rcp26_20)
plot(tiff_file_2b_soilmoist_rcp26_20)  

#stacking raster files
#Convert SpatRaster to RasterLayer
#tiff_file_2b_soilmoist_rcp26_20_raster <- raster(tiff_file_2b_soilmoist_rcp26_20)

#Stack the rasters
#rs_rcp26_20 <- stack(tiff_file_2b_noirr_rcp26_20, tiff_file_2b_soilmoist_rcp26_20_raster)


rs_rcp26_20 <- c(tiff_file_2b_noirr_rcp26_20, tiff_file_2b_soilmoist_rcp26_20)

#Plot the raster stack
plot(rs_rcp26_20)



```
Extract the raster values for each layer and combine them into a dataset for modeling. 
```{r}
#Convert raster layers to a data frame
noirr_rcp26_20 <- as.data.frame(tiff_file_2b_noirr_rcp26_20, xy = TRUE, na.rm = TRUE)

soilmoist_rcp26_20 <- as.data.frame(tiff_file_2b_soilmoist_rcp26_20, xy = TRUE, na.rm = TRUE)

#Inspect data
head(noirr_rcp26_20)
head(soilmoist_rcp26_20)
```

```{r}
#Load food security data
Food_security_values <- read.csv("Socio-economic-survey-ETH_2020_FSI.csv")

#Check 'lon' and 'lat' columns 
head(Food_security_values)

#Convert to SpatVector with correct CRS
Food_security_locations <- vect(Food_security_values, geom = c("lon", "lat"), crs = "EPSG:4326")

Food_security_locations <- project(Food_security_locations, crs(rs_rcp26_20))

#Extract raster values at the given locations
Food_Security_Extract1 <- terra::extract(rs_rcp26_20, Food_security_locations)

#View extracted values
head(Food_Security_Extract1)

#Extract raster values at the point locations
Food_Security_Extract1 <- raster::extract(rs_rcp26_20, Food_security_locations)

#Combine with the point coordinates
Food_Security_Extract1_df <- cbind(as.data.frame(Food_security_locations), Food_Security_Extract1)

#View 
head(Food_Security_Extract1_df)
print(Food_Security_Extract1_df)


```


```{r}
#Plot raster
plot(rs_rcp26_20[[1]], main = "Raster and Points")

points(Food_security_locations, col = "red", pch = 19)
```

```{r}
#Load Ethiopia shapefile
ethiopia_map <- st_read("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ML Data/TIFF Rename/gadm41_ETH_0.shp")  

#Plot raster 
plot(ethiopia_map[[1]], main = "Raster and Points")

points(Food_security_locations, col = "red", pch = 19)
```



```{r}


#Load Ethiopia boundary
ethiopia <- ne_countries(scale = "medium", country = "Ethiopia", returnclass = "sf")

#Convert Food Security Points to an sf object
Food_security_points <- st_as_sf(Food_security_values, coords = c("lon", "lat"), crs = 4326)

#Change FSI_Bool is a factor for color mapping
Food_security_points$FSI_Bool <- as.factor(Food_security_values$FSI_Bool)

#Extract coordinates into separate columns
Food_security_points$longitude <- st_coordinates(Food_security_points)[, 1]
Food_security_points$latitude <- st_coordinates(Food_security_points)[, 2]

#Plot Ethiopia with points colored by presence/absence
ggplot() +
  geom_sf(data = ethiopia, fill = "gray90", color = "black") + 
  geom_point(data = Food_security_points, aes(x = longitude, y = latitude, color = FSI_Bool), size = 2, alpha = 0.7) +
  scale_color_manual(values = c("red", "blue"), labels = c("Absence (0)", "Presence (1)")) +  
  theme_minimal() +
  labs(title = "Presence and Absence Points in Ethiopia",
       color = "FSI Presence",
       x = "Longitude", y = "Latitude")
```


```{r}
#Standardise column names
names(Food_Security_Variables_Values_rcp26_20) <- make.names(names(Food_Security_Variables_Values_rcp26_20), unique = TRUE)

#Check cleaned column names
colnames(Food_Security_Variables_Values_rcp26_20)
```


Split the data into training and testing subsets and train model with FSI
```{r}
#Split data into training and testing sets
set.seed(123)

#Select the relevant features for modeling
raster_values <- Food_Security_Variables_Values_rcp26_20 %>%
  select(Ethiopia_wheat_RCP26_noirrigation_5km_2020.copy, 
         CLM45_RPC_26_Weighted_Average_Soil_Moisture_5km_2020.copy, 
         FSI) %>%
  na.omit() 

colnames(raster_values) <- c("Wheat_Index", "Soil_Moisture", "FSI")

#Split the data
train_index <- sample(1:nrow(raster_values), size = 0.7 * nrow(raster_values))
train_data <- raster_values[train_index, ]
test_data <- raster_values[-train_index, ]

#Convert FSI to a factor
raster_values$FSI <- as.factor(raster_values$FSI)
train_data$FSI <- factor(train_data$FSI, levels = levels(raster_values$FSI))
test_data$FSI <- factor(test_data$FSI, levels = levels(raster_values$FSI))

#Train a random forest model
rf_model <- randomForest(FSI ~ ., data = train_data, ntree = 100)

#Evaluate the model
predictions <- predict(rf_model, test_data)

test_data$FSI <- as.factor(test_data$FSI)
predictions <- as.factor(predictions)

#Confusion matrix for model evaluation
confusion_matrix <- confusionMatrix(predictions, test_data$FSI)
print(confusion_matrix)
```

```{r}

#Combine test data with predictions
test_data_with_preds <- test_data
test_data_with_preds$Predicted <- predictions

#Create a heatmap of predicted vs actual FSI
ggplot(test_data_with_preds, aes(x = FSI, y = Predicted)) +
  geom_bin2d() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(
    title = "Heatmap of Predictions vs Actual (Original Data)",
    x = "Actual FSI",
    y = "Predicted FSI"
  ) +
  theme_minimal()



#Plot confusion matrix
conf_matrix_df <- as.data.frame(confusion_matrix$table)
ggplot(conf_matrix_df, aes(Reference, Prediction, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), color = "white", size = 5) +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(
    title = "Confusion Matrix (Original Data)",
    x = "Actual Class",
    y = "Predicted Class"
  ) +
  theme_minimal()
```

Split the data into training and testing subsets and train model with FSI_Bool
```{r}

set.seed(123)

#Select relevant features
raster_values1 <- Food_Security_Variables_Values_rcp26_20 %>%
  select(Ethiopia_wheat_RCP26_noirrigation_5km_2020.copy, 
         CLM45_RPC_26_Weighted_Average_Soil_Moisture_5km_2020.copy, 
         FSI_Bool) %>%
  na.omit()

colnames(raster_values1) <- c("Wheat_Index", "Soil_Moisture", "FSI_Bool")

#Check FSI_Bool is a factor
raster_values1$FSI_Bool <- as.factor(raster_values1$FSI_Bool)

#Split data into training and testing sets
train_index1 <- sample(1:nrow(raster_values1), size = 0.7 * nrow(raster_values1))
train_data1 <- raster_values1[train_index1, ]
test_data1 <- raster_values1[-train_index1, ]

test_data1$FSI_Bool <- factor(test_data1$FSI_Bool, levels = levels(train_data1$FSI_Bool))

#Train the random forest model
rf_model1 <- randomForest(FSI_Bool ~ ., data = train_data1, ntree = 100)

#Make predictions
predictions1 <- predict(rf_model1, test_data1)

#Convert predictions to factor with same levels as FSI_Bool in test data
predictions1 <- factor(predictions1, levels = levels(test_data1$FSI_Bool))

#Confusion matrix for model evaluation
confusion_matrix <- confusionMatrix(predictions1, test_data1$FSI_Bool)

#Print the confusion matrix
print(confusion_matrix)
```

```{r}

set.seed(123)

#Select the relevant features for modeling
raster_values1 <- Food_Security_Variables_Values_rcp26_20 %>%
  select(Ethiopia_wheat_RCP26_noirrigation_5km_2020.copy, 
         CLM45_RPC_26_Weighted_Average_Soil_Moisture_5km_2020.copy, 
         FSI_Bool)

colnames(raster_values1) <- c("Wheat_Index", "Soil_Moisture", "FSI_Bool")

#Convert FSI_Bool to a boolean variable where 0 = food insecurity absent, 1 = food security present
raster_values1$FSI_Bool <- ifelse(raster_values1$FSI_Bool == 0, "Absent", "Present")
raster_values1$FSI_Bool <- as.factor(raster_values1$FSI_Bool) 

#Split the data into training and testing sets
train_index1 <- sample(1:nrow(raster_values1), size = 0.7 * nrow(raster_values1))
train_data1 <- raster_values1[train_index1, ]
test_data1 <- raster_values1[-train_index1, ]

#Train a random forest model
rf_model1 <- randomForest(FSI_Bool ~ ., data = train_data1, ntree = 100)

#Evaluate the model
predictions1 <- predict(rf_model1, test_data1)

#Confusion matrix for model evaluation
confusion_matrix1 <- confusionMatrix(predictions1, test_data1$FSI_Bool)
print(confusion_matrix1)
```

```{r}
#Combine test data with predictions
test_data_with_preds <- test_data1
test_data_with_preds$Predicted <- predictions1

#Create a heatmap of predicted vs actual FSI_Bool
ggplot(test_data_with_preds, aes(x = FSI_Bool, y = Predicted)) +
  geom_bin2d() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(
    title = "Heatmap of Predictions vs Actual (Original Data)",
    x = "Actual FSI_Bool",
    y = "Predicted FSI_Bool"
  ) +
  theme_minimal()


#Plot confusion matrix
conf_matrix_df1 <- as.data.frame(confusion_matrix1$table)
ggplot(conf_matrix_df1, aes(Reference, Prediction, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), color = "white", size = 5) +
  scale_fill_gradient(low = "blue", high = "red") +
  labs(
    title = "Confusion Matrix (Original Data)",
    x = "Actual Class",
    y = "Predicted Class"
  ) +
  theme_minimal()
```


Generalised additive model
```{r}

raster_values1$FSI_Bool <- as.factor(raster_values1$FSI_Bool)

set.seed(123) 
train_index1 <- sample(1:nrow(raster_values1), size = 0.7 * nrow(raster_values1))
train_data1 <- raster_values1[train_index1, ]
test_data1 <- raster_values1[-train_index1, ]

#Fit GAM with smoothing splines on Wheat_Index and Soil_Moisture
gam_model1 <- gam(FSI_Bool ~ s(Wheat_Index) + s(Soil_Moisture), 
                  family = binomial, 
                  data = train_data1)

#Print model summary
summary(gam_model1)

```

Dynamic Generalised Additive model:

```{r}
#Ensure FSI_Bool is a factor (0 = Absent, 1 = Present)
raster_values1$FSI_Bool <- as.factor(raster_values1$FSI_Bool)

#Check data structure
str(raster_values1)

#Split into training (70%) and test (30%) sets
set.seed(123) 
train_index1 <- sample(1:nrow(raster_values1), size = 0.7 * nrow(raster_values1))
train_data1 <- raster_values1[train_index1, ]
test_data1 <- raster_values1[-train_index1, ]

#Fit DGAMLSS using smooth functions for predictors
dgamlss_model <- gamlss(FSI_Bool ~ pb(Wheat_Index) + pb(Soil_Moisture), 
                        family = BI(),
                        data = train_data1)

#Print model summary
summary(dgamlss_model)

#Plot the effects of Wheat Yield and Soil Moisture
par(mfrow = c(1, 2))  
term.plot(dgamlss_model, pages = 1)

#Predict probabilities on test set
test_data1$predicted_prob <- predict(dgamlss_model, newdata = test_data1, type = "response")

#Convert probabilities to binary classification (threshold = 0.5)
test_data1$predicted_class <- ifelse(test_data1$predicted_prob > 0.5, "Present", "Absent")

#Convert to factors for classification evaluation
test_data1$predicted_class <- factor(test_data1$predicted_class, levels = levels(test_data1$FSI_Bool))

#Confusion Matrix
confusion_matrix_dgamlss <- confusionMatrix(test_data1$predicted_class, test_data1$FSI_Bool)

#Print Confusion Matrix
print(confusion_matrix_dgamlss)

ggplot(test_data1, aes(x = Wheat_Index, y = predicted_prob, color = FSI_Bool)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess") +
  labs(title = "Predicted Probability of Food Security",
       x = "Wheat Index",
       y = "Predicted Probability")


```

```{r}
#confusion matrix
conf_matrix <- confusionMatrix(test_data1$predicted_class, test_data1$FSI_Bool)

#Extract the confusion matrix as a data frame
conf_matrix_table <- as.data.frame(conf_matrix$table)

#Rename columns for better readability
colnames(conf_matrix_table) <- c("Actual", "Predicted", "Frequency")

#Print the confusion matrix table
print(conf_matrix_table)

#Compute the confusion matrix
conf_matrix <- confusionMatrix(test_data1$predicted_class, test_data1$FSI_Bool)

#Print detailed confusion matrix results
print(conf_matrix)

#Plot confusion matrix
ggplot(conf_matrix_table, aes(x = Actual, y = Predicted, fill = Frequency)) +
  geom_tile(color = "black") + 
  geom_text(aes(label = Frequency), color = "white", size = 5) + 
  scale_fill_gradient(low = "lightblue", high = "darkblue") + 
  theme_minimal() +
  labs(
    title = "Confusion Matrix",
    x = "Actual Class",
    y = "Predicted Class",
    fill = "Count"
  ) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 14))
```

Maximum Entropy:
```{r}
#Reload the food security dataset
Food_security_values <- read.csv("Socio-economic-survey-ETH_2020_FSI.csv")

#Check FSI_Bool is binary
Food_security_values$FSI_Bool <- as.numeric(Food_security_values$FSI_Bool)

Food_security_values <- na.omit(Food_security_values)

#Merge Food Security Data with Raster Data
raster_values1 <- merge(Food_security_values, raster_values1, by = c("lon", "lat"), all.y = TRUE)

#Remove rows where FSI_Bool is still NA
raster_values1 <- raster_values1 %>% filter(!is.na(FSI_Bool))

#Split into training (70%) and test (30%) sets
set.seed(123)  
train_index1 <- sample(1:nrow(raster_values1), size = 0.7 * nrow(raster_values1))
train_data1 <- raster_values1[train_index1, ]
test_data1 <- raster_values1[-train_index1, ]

#Train Maximum Entropy Model (MaxEnt)
maxent_model <- maxnet(
  p = train_data1$FSI_Bool,
  data = train_data1 %>% select(Wheat_Index, Soil_Moisture), 
  f = maxnet.formula(train_data1$FSI_Bool, train_data1 %>% select(Wheat_Index, Soil_Moisture))
)

#Print
summary(maxent_model)


```

```{r}
#Ensure FSI_Bool is numeric (0 = Absent, 1 = Present)
raster_values1$FSI_Bool <- as.numeric(as.character(raster_values1$FSI_Bool))

#Split into training (70%) and test (30%) sets
set.seed(123)  
train_index1 <- sample(1:nrow(raster_values1), size = 0.7 * nrow(raster_values1))
train_data1 <- raster_values1[train_index1, ]
test_data1 <- raster_values1[-train_index1, ]

#Train Maximum Entropy Model (MaxEnt)
maxent_model <- maxnet(
  p = train_data1$FSI_Bool,  
  data = train_data1 %>% select(Wheat_Index, Soil_Moisture), 
  f = maxnet.formula(train_data1$FSI_Bool, train_data1 %>% select(Wheat_Index, Soil_Moisture))
)

#Print 
summary(maxent_model)

#Predict probabilities on test data
test_data1$predicted_prob <- predict(maxent_model, test_data1 %>% select(Wheat_Index, Soil_Moisture), type = "exponential")

#Convert probabilities to binary classification (threshold = 0.5)
test_data1$predicted_class <- ifelse(test_data1$predicted_prob > 0.5, "Present", "Absent")

#Convert to factors for classification evaluation
test_data1$predicted_class <- factor(test_data1$predicted_class, levels = levels(factor(test_data1$FSI_Bool)))

#Compute Confusion Matrix
confusion_matrix_maxent <- confusionMatrix(test_data1$predicted_class, factor(test_data1$FSI_Bool, levels = c("Absent", "Present")))

#Print Confusion Matrix
print(confusion_matrix_maxent)

#Convert confusion matrix to data frame
conf_matrix_table <- as.data.frame(confusion_matrix_maxent$table)

colnames(conf_matrix_table) <- c("Actual", "Predicted", "Frequency")

#Plot confusion matrix heatmap
ggplot(conf_matrix_table, aes(x = Actual, y = Predicted, fill = Frequency)) +
  geom_tile(color = "black") +  
  geom_text(aes(label = Frequency), color = "white", size = 6) +  
  scale_fill_gradient(low = "lightblue", high = "darkblue") +  
  theme_minimal() +
  labs(
    title = "MaxEnt Model - Confusion Matrix",
    x = "Actual Class",
    y = "Predicted Class",
    fill = "Count"
  ) +
  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(hjust = 0.5, size = 14))

#Extract feature importance
importance <- as.data.frame(varImp(maxent_model))

#Plot feature importance
ggplot(importance, aes(x = reorder(rownames(importance), Overall), y = Overall, fill = Overall)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Feature Importance (MaxEnt Model)", x = "Features", y = "Importance Score") +
  scale_fill_gradient(low = "lightblue", high = "darkblue")

#Extract key evaluation metrics
accuracy <- confusion_matrix_maxent$overall["Accuracy"]
kappa <- confusion_matrix_maxent$overall["Kappa"]
sensitivity <- confusion_matrix_maxent$byClass["Sensitivity"]
specificity <- confusion_matrix_maxent$byClass["Specificity"]
precision <- confusion_matrix_maxent$byClass["Pos Pred Value"]
f1_score <- 2 * ((precision * sensitivity) / (precision + sensitivity))

#Print results
cat("Accuracy:", accuracy, "\n")
cat("Kappa:", kappa, "\n")
cat("Sensitivity (Recall):", sensitivity, "\n")
cat("Specificity:", specificity, "\n")
cat("Precision:", precision, "\n")
cat("F1-Score:", f1_score, "\n")


```


```{r}
#Load Ethiopia shapefile 
ethiopia_map <- st_read("/Users/maiawatson-hearne/Desktop/UCL Geography Y3/Dissertation/Data/ISIMPI/gadm41_ETH_0.shp")  

#Ensure the coordinate reference system (CRS) matches
ethiopia_map <- st_transform(ethiopia_map, crs = 4326)
```


After training, use model to predict labels for the entire raster dataset
```{r}
#Include latitude and longitude columns 
raster_values1 <- Food_Security_Variables_Values_rcp26_20 %>%
  select(
    Ethiopia_wheat_RCP26_noirrigation_5km_2020.copy,
    CLM45_RPC_26_Weighted_Average_Soil_Moisture_5km_2020.copy,
    lat,
    lon,
    FSI_Bool
  ) 

test_data_with_preds$lat <- raster_values1[-train_index1, "lat"]
test_data_with_preds$lon <- raster_values1[-train_index1, "lon"]
view(test_data_with_preds)

#Convert test data with predictions to an sf object
test_data_sf <- st_as_sf(test_data_with_preds, coords = c("lon", "lat"), crs = 4326)

#Plot the map with predictions
ggplot() +
  geom_sf(data = ethiopia_map, fill = "gray90", color = "black") +  
  geom_sf(data = test_data_sf, aes(color = Predicted), size = 1, alpha = 0.7) +  
  scale_color_manual(values = c("Absent" = "red", "Present" = "green")) + 
  labs(
    title = "Predicted Food Security in Ethiopia",
    color = "Predicted FSI_Bool"
  ) +
  theme_minimal()
```


After training, use model to predict labels for the entire raster dataset
```{r}
#Predict on raster layers
predicted_raster <- predict(tiff_file, rf_model)

#Visualize the predicted raster
plot(predicted_raster)
```

Maximum entropy
```{r}
#Prepare raster_values1 dataset with predictors
raster_values2 <- Food_Security_Variables_Values_rcp26_20 %>%
  select(
    Ethiopia_wheat_RCP26_noirrigation_5km_2020.copy,
    CLM45_RPC_26_Weighted_Average_Soil_Moisture_5km_2020.copy,
    lat,
    lon,
    FSI_Bool
  ) 

colnames(raster_values2) <- c("Wheat_Index", "Soil_Moisture", "lat", "lon", "FSI_Bool")

#Convert FSI_Bool to binary (0 = absent, 1 = present)
raster_values2$FSI_Bool <- as.numeric(raster_values2$FSI_Bool)
```

```{r}
#Ensure FSI_Bool is numeric (0 = absent, 1 = present)
train_data$FSI_Bool <- as.numeric(as.character(train_data$FSI_Bool))
test_data$FSI_Bool <- as.numeric(as.character(test_data$FSI_Bool))

#Train the MaxEnt model
maxent_model <- maxnet(
  p = train_data$FSI_Bool,  
  data = train_data[, c("Wheat_Index", "Soil_Moisture")],  
  f = maxnet.formula(
    p = train_data$FSI_Bool,
    data = train_data[, c("Wheat_Index", "Soil_Moisture")]
  )
)
```

```{r}
#Predict probabilities on the test dataset
test_data$Predicted_Prob <- predict(maxent_model, test_data[, c("Wheat_Index", "Soil_Moisture")], type = "logistic")
```

```{r}
#Convert probabilities to binary predictions using a threshold (e.g., 0.5)
test_data$Predicted_Binary <- ifelse(test_data$Predicted_Prob > 0.5, 1, 0)

#Confusion matrix
confusion_matrix <- confusionMatrix(factor(test_data$Predicted_Binary),
                                    factor(test_data$FSI_Bool))
print(confusion_matrix)
```


Trialing Dynamic Generalized Additive Models

```{r}
set.seed(123)
n <- 1000
data <- data.frame(
  Time = 1:n,
  Wheat_Index = runif(n, 50, 100),
  Soil_Moisture = runif(n, 20, 80),
  FSI_Bool = as.factor(rbinom(n, 1, 0.5))
)

#Convert to factor
data$FSI_Bool <- factor(data$FSI_Bool, levels = c(0, 1), labels = c("absent", "present"))
```

```{r}

#Fit the DGAM
model_dgam <- gam(
  FSI_Bool ~ s(Wheat_Index) + s(Soil_Moisture) + s(Time, bs = "cr"),
  data = data,
  family = binomial(link = "logit")
)

#Summarise the model
summary(model_dgam)
```
```{r}
#Plot smooth terms
plot(model_dgam, pages = 1, seWithMean = TRUE, rug = TRUE)
```

```{r}
#Generate new data for prediction
new_data <- data.frame(
  Time = seq(1, 200, length.out = 50),
  Wheat_Index = runif(50, 50, 100),
  Soil_Moisture = runif(50, 20, 80)
)

#Predict probabilities
predictions <- predict(model_dgam, newdata = new_data, type = "response")

#Add predictions to new_data
new_data$Predicted_Prob <- predictions

head(new_data)
```

```{r}

#Predict on test data
pred_probs <- predict(model_dgam, newdata = data, type = "response")
pred_classes <- ifelse(pred_probs > 0.5, "present", "absent")

#Confusion Matrix
confusion_matrix <- table(Predicted = pred_classes, Actual = data$FSI_Bool)
print(confusion_matrix)

#ROC Curve
roc_curve <- roc(data$FSI_Bool, pred_probs, levels = c("absent", "present"))
plot(roc_curve, main = "ROC Curve")
auc(roc_curve)  
```

```{r}
#Create a distributed lag structure for Wheat_Index
lag_wheat <- crossbasis(data$Wheat_Index, lag = 5, argvar = list(fun = "ns", df = 3), arglag = list(fun = "ns", df = 3))

#Fit the GAM with lag structure
model_lag_gam <- gam(
  FSI_Bool ~ s(Soil_Moisture) + s(Time, bs = "cr") + lag_wheat,
  data = data,
  family = binomial(link = "logit")
)

summary(model_lag_gam)
```

